Contextuality and the Born rule in fuzzy quantum theories

### 2017-04-08 prop_letter.tex
prop_letter.tex Updating here and there...

Although considering IVPM is new for a quantum theory, people have
already defined classical IVPMs~\cite{JamisonLodwick2004}, which
are essentially the same as quantum IVPMs on a restricted mutually
commuting sub-event space~$\events'\subseteq\events$, where any
subset of $\events$ is called a sub-event space if it satisfies the
follow conditions. 
\begin{itemize}
\item $\mathbb{0}\in\events'$. 
\item $\mathbb{1}\in\events'$. 
\item For any projection~$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For a set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$. 
\end{itemize}
Quantum IVPM not only stems from classical IVPM but also generalizes
quantum real-valued probability measures. Consider the collection
of all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
and a function~$\iota_{\infty}:\left[0,1\right]\rightarrow\mathscr{I}_{\infty}$
defined by $\iota_{\infty}\left(x\right)=\left[x,x\right]$. Then,
given a quantum real-valued probability measure~$\mu:\events\rightarrow\left[0,1\right]$,
$\bar{\mu}(P)=\iota_{\infty}\left(\mu(P)\right)$ is a quantum IVPM,
and vice versa. In another word, quantum IVPMs mapping to $\mathscr{I}_{\infty}$
are just the same as quantum real-valued probability measures where
Gleason's theorem can apply.

@@

Recall when providing infinity measurement resource, the quantum probability
measures are simply induced by the Born rule for $d\ge3$. If we increase
the measurement resource, i.e., pick finer intervals, it is natural
to ask whether quantum IVPMs approach the Born rule. In order to understand
this question, it is instructive to consider the 2-dimensional non-Born
quantum probability measure~$\mu_{0}:\events\rightarrow[0,1]$ defined
as follow:
\begin{equation}
\mu_{0}(P)=\begin{cases}
\mu_{\ket{\ps}}^{\mathrm{B}}(P) & \textrm{, if }P\in\events_{0}\textrm{ ;}\\
\mu_{\ket{0}}^{\mathrm{B}}(P) & \textrm{, otherwis,}
\end{cases}
\end{equation}
where $\events_{0}=\left\{ \mathbb{0},\proj{\ps},\proj{\ms},\mathbb{1}\right\} $
is a sub-event space. In another word, $\mu_{0}$ is consistent with
the Born rule on every proper sub-event space, but not on the set
of all projectors.

Given a quantum IVPM~$\bar{\mu}$, the preceding idea motivates the
definition of \emph{core} relative to a sub-event space~$\events'\subseteq\events$
as follow.\footnote{\yutsung{{I changed the definition of core} to density matrix. While
this change might make it easier to explain the core of quantum IVPMs,
it could increase the difficulty to explain classical IVPMs because
we might need to explain the meaning of density matrix and the Born
rule for classical probability measures...}}
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in\events'.~\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
Although the elements of a core are density matrices, we may just
use the state vectors for pure states if it is clear according to
the context. For example, $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events_{0}\right)=\left\{ \ket{\ps}\right\} $
and $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events\right)=\emptyset$,
where $\circ$ denote function composition.

@@

Let $\events_{1}$ be the minimal event space containing both $\proj{0}$
and $\proj{\ps}$. Because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
for any mixed state~$\rho\in\mathrm{core}\left(\bar{\mu},\events_{1}\right)$,
we have $\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=0$.
Hence, $\rho$ is actually the pure state orthogonal to both $\ket{0}$
and $\ket{\ps}$. The only choice of $\rho$ is $\ket{2}$, and it
is easy to verify $\ket{2}$ is actually in $\mathrm{core}\left(\bar{\mu},\events_{1}\right)$.

Similarly, let $\events_{2}$ be the minimal event space containing
both $\proj{0}$ and $\proj{\ps'}$. We have $\mathrm{core}\left(\bar{\mu},\events_{2}\right)=\left\{ \ket{1}\right\} $.
Therefore, 
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events\right)\subseteq\mathrm{core}\left(\bar{\mu},\events_{1}\right)\cap\mathrm{core}\left(\bar{\mu},\events_{2}\right)=\emptyset\textrm{ ,}
\end{equation}
and must be empty.\end{example}

>>>>

Although considering IVPM is new for a quantum theory, people have
already defined classical IVPMs~\cite{JamisonLodwick2004}, which
are essentially the same as quantum IVPMs on a restricted mutually
commuting sub-event space~$\events'\subseteq\events$, where any
subset of $\events$ is called a sub-event space if it satisfies the
follow conditions. 
\begin{itemize}
\item $\mathbb{0}\in\events'$. 
\item $\mathbb{1}\in\events'$. 
\item For any projection~$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For a set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$. 
\end{itemize}
Quantum IVPM not only stems from classical IVPM but also generalizes
quantum real-valued probability measures. Consider the collection
of all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
and a function~$\iota_{\infty}:\left[0,1\right]\rightarrow\mathscr{I}_{\infty}$
defined by 
\begin{equation}
\iota_{\infty}\left(x\right)=\left[x,x\right]\textrm{ .}\label{eq:iota_infty}
\end{equation}
Then, given a quantum real-valued probability measure~$\mu:\events\rightarrow\left[0,1\right]$,
$\bar{\mu}(P)=\iota_{\infty}\left(\mu(P)\right)$ is a quantum IVPM,
and vice versa. In another word, quantum IVPMs mapping to $\mathscr{I}_{\infty}$
are just the same as quantum real-valued probability measures where
Gleason's theorem can apply.

@@

Formally, we can ask: what can we deduce about the state of a quantum
system given a quantum interval-valued probability measure, i.e.,
given observations done with finite resources. In the case that the
intervals are infinitely precise, the question reduces to Gleason's
theorem which states that the state of the quantum system is uniquely
determined by the probability measure. But surely the less resources
are available, the less precise the intervals, and the less we expect
to know about the state of the system.

In order to understand this question, it is instructive to consider
the 2-dimensional quantum probability measure~$\mu_{0}:\events\rightarrow[0,1]$
defined as follow:
\begin{equation}
\mu_{0}(P)=\begin{cases}
\mu_{\ket{\ps}}^{\mathrm{B}}(P) & \textrm{, if }P\in\events_{0}\textrm{ ;}\\
\mu_{\ket{0}}^{\mathrm{B}}(P) & \textrm{, otherwis,}
\end{cases}
\end{equation}
where $\events_{0}=\left\{ \mathbb{0},\proj{\ps},\proj{\ms},\mathbb{1}\right\} $
is a sub-event space. $\mu_{0}$ is not induced by the Born rule as
a whole, but is consistent with the Born rule on $\events_{0}$ and
some other proper sub-event spaces. Formally, given a quantum IVPM~$\bar{\mu}$,
the \emph{core} relative to a sub-event space~$\events'\subseteq\events$
as follow.\footnote{\yutsung{I changed the definition of core to embedding the Born rule.
While this change might make it easier to explain the core of quantum
IVPMs, it could increase the difficulty to explain classical IVPMs
because we might need to explain the meaning of the Born rule for
classical probability measures...}}
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in\events'.~\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
For example, $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events_{0}\right)=\left\{ \proj{\ps}\right\} $
and $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events\right)=\emptyset$,
where $\circ$ denote function composition, and $\iota_{\infty}$
is defined in Eq.~(\ref{eq:iota_infty}).

@@

Let $\events_{1}$ be the minimal event space containing both $\proj{0}$
and $\proj{\ps}$. Because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
for any mixed state~$\rho\in\mathrm{core}\left(\bar{\mu},\events_{1}\right)$,
we have $\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=0$.
Hence, $\rho$ is actually the pure state orthogonal to both $\ket{0}$
and $\ket{\ps}$. The only choice of $\rho$ is $\ket{2}$, and it
is easy to verify $\proj{2}$ is actually in $\mathrm{core}\left(\bar{\mu},\events_{1}\right)$.

Similarly, let $\events_{2}$ be the minimal event space containing
both $\proj{0}$ and $\proj{\ps'}$. We have $\mathrm{core}\left(\bar{\mu},\events_{2}\right)=\left\{ \proj{1}\right\} $.
Therefore, 
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events\right)\subseteq\mathrm{core}\left(\bar{\mu},\events_{1}\right)\cap\mathrm{core}\left(\bar{\mu},\events_{2}\right)=\emptyset\textrm{ ,}
\end{equation}
and must be empty.\end{example}
